---
title: "DDOT project evaluations"
author: "pete rodrigue"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8)

# Libraries we need:
library(readr)
library(ggplot2)
library(dplyr)
library(stringr)
library(sf)
library(leaflet)

# google sheet w/ the data:
url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vQtDsTt9VVyn-wptsMT6vxtCM2J1dPYWji90OOWj1v4c4pPz6Kwa6mQtOk096EU3npvJ9g4fcjnwLpp/pub?gid=1445637369&single=true&output=csv"

# load the data from the google sheet:
df <- read_csv(url, show_col_types = F)
```


```{r}
to_plot <-
  df %>%
  filter(Outcome %in% c("Injury crashes", "All crashes")) %>%
  mutate(Change = Change*-1)

ggplot(to_plot) +
  geom_bar(aes(x=`Project name`, y=Change, fill=Outcome), stat="identity", position="dodge") +
  labs(y="% decrease in crashes", x="") +
  ggtitle("All projects saw decreases in crashes") +
  geom_hline(aes(yintercept=mean(Change[Outcome == "All crashes"], na.rm=T)), color="#F8766D", linetype="dashed", linewidth=.8, alpha=.5) +
  geom_hline(aes(yintercept=mean(Change[Outcome == "Injury crashes"], na.rm=T)), color="#00BFC4", linetype="dashed", linewidth=.8, alpha=.5) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

```{r}
to_plot <-
  df %>%
  filter(Outcome %in% c("Pedestrian injury crashes")) %>%
  mutate(Change = Change*-1)

ggplot(to_plot) +
  geom_bar(aes(x=`Project name`, y=Change), stat="identity", fill="seagreen") +
  labs(y="% decrease in crashes", x="") +
  ggtitle("Pedestrian crashes fell after project completion, sometimes to zero") +
  geom_hline(aes(yintercept=mean(Change, na.rm=T)), color="grey", linetype="dashed", size=1) +
  theme_minimal() +
  theme(legend.title = element_blank())
```



```{r}
to_plot <-
  df %>%
  filter(str_detect(Outcome, "Private motor vehicle")) %>%
  group_by(`Project name`, Outcome) %>%
  summarize(change = mean(Change, na.rm=T)) %>%
  ungroup() %>%
  mutate(Outcome = ifelse(Outcome == "Private motor vehicle AM peak travel time (seconds)", 
                          "AM traffic travel time",
                          "PM traffic travel time"))

ggplot(to_plot) +
  geom_point(
    aes(x=change, color=Outcome, y=`Project name`), size=3, alpha=.7
  ) +
  geom_vline(aes(xintercept=0), color="black") +
  geom_vline(aes(xintercept=mean(change[Outcome == "AM traffic travel time"], na.rm=T)), color="#F8766D", linetype="dashed") +
  geom_vline(aes(xintercept=mean(change[Outcome == "PM traffic travel time"], na.rm=T)), color="#00BFC4", linetype="dashed") +
  labs(x="Change in travel time in seconds", y="") +
  theme_minimal() +
  xlim(c(min(to_plot$change)-10, max(to_plot$change)+25)) +
  theme(legend.title = element_blank()) +
  ggtitle("Drivers' travel times faster on average after project completion")
```


```{r}
bl <- st_read(
  "https://opendata.dc.gov/api/download/v1/items/482ed4051575418caaae182a35c23cd7/geojson?layers=2"
) %>%
  # filter(
  #   (!is.na(BIKELANE_PROTECTED)) |
  #     (!is.na(BIKELANE_BUFFERED))
  #   ) %>%
   st_zm(drop = TRUE, what = "ZM") %>%  # drops Z (and M if present) → pure XY
  # Leaflet requires WGS84 (EPSG:4326) — reproject if needed
  st_transform(4326) %>%
  st_cast("MULTILINESTRING") 

pal <- colorNumeric(
  palette = "YlOrRd", 
  domain = bl$build_year
)

leaflet(bl) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolylines(
    color            = ~pal(build_year),
    weight           = 4,
    opacity          = 0.85,
    highlightOptions = highlightOptions(  
      weight       = 6,
      color        = "#222222",
      opacity      = 1,
      bringToFront = TRUE
    ),
    label        = ~paste0("<b>", ROUTENAME, "</b><br>build yr: ", build_year) |>
                     lapply(htmltools::HTML),
    labelOptions = labelOptions(
      textsize  = "12px",                     
      direction = "auto"
    )
  )
```


```{r}
bl <- 
  st_read("bike_lanes_labeled.geojson") %>%
     st_zm(drop = TRUE, what = "ZM") %>%  # drops Z (and M if present) → pure XY
  # Leaflet requires WGS84 (EPSG:4326) — reproject if needed
  st_transform(4326) %>%
  st_cast("MULTILINESTRING") 
```

```{r}
# load crash data
# data downloaded from here:
# https://opendata.dc.gov/datasets/crashes-in-dc/about
cd <- readr::read_csv("Crashes_in_DC.csv")

# create year and month of crash variables:
cd$year  <- as.numeric(substr(x = cd$REPORTDATE, start = 1, stop = 4))
cd$month <- as.numeric(substr(x = cd$REPORTDATE, start = 6, stop = 7))

# remove rows with no date information:
cd <- filter(cd, !is.na(year))

# project crash data:
cd <- cd |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

injury_vars <- c(
  "MAJORINJURIES_BICYCLIST",   "MINORINJURIES_BICYCLIST",
  "UNKNOWNINJURIES_BICYCLIST", "FATAL_BICYCLIST",
  "MAJORINJURIES_DRIVER",      "MINORINJURIES_DRIVER",
  "UNKNOWNINJURIES_DRIVER",    "FATAL_DRIVER",
  "MAJORINJURIES_PEDESTRIAN",  "MINORINJURIES_PEDESTRIAN",
  "UNKNOWNINJURIES_PEDESTRIAN","FATAL_PEDESTRIAN"
)

cd <- select(cd, CRIMEID, year, month, injury_vars)
```


```{r}
# look just at PBLs
pbl <- filter(bl, !is.na(build_year))
```

```{r}
# ── 1. Unify segments by ROUTENAME, keeping max and min build_year ──────────────────
pbl_union <- pbl |>
  # fix route naming, combine PBLs that were adjacent and installed in the same year
  mutate(ROUTENAME = if_else(ROUTENAME=="Ramp-36001642", "IRVING ST NW", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="IRVING ST NE", "IRVING ST NW", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="TUNLAW RD NW", "NEW MEXICO AVE NW", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="15TH ST SW", "15TH ST NW", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="FLORIDA AVE NW", "9TH ST NW", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="7TH ST NW", "WARDER ST NW", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="5TH ST NW", "PARK PL NW", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="4TH ST NW", "PARK PL NW", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="6TH ST NE", "BRENTWOOD PKWY NE", ROUTENAME)) %>%
  mutate(ROUTENAME = if_else(ROUTENAME=="M st SE custom", "M ST SE", ROUTENAME)) %>%
  # mutate(ROUTENAME = if_else(ROUTENAME=="", "", ROUTENAME)) %>%
  group_by(ROUTENAME) |>
  summarise(
    max_build_year = max(build_year, na.rm = TRUE),
    min_build_year = min(build_year, na.rm = TRUE),
    .groups    = "drop"
  )
# summarise() on an sf object auto-unions the geometries within each group

## visual check:
leaflet(bl) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolylines(
    color            = "green",
    weight           = 6,
    opacity          = 0.85,
    highlightOptions = highlightOptions(
      weight       = 6,
      color        = "#222222",
      opacity      = 1,
      bringToFront = TRUE
    ),
    label        = ~paste0("<b>", ROUTENAME#,
                           # "</b><br>min build yr: ", min_build_year,
                           # "<br>max build yr: ", max_build_year
                           ) |>
                     lapply(htmltools::HTML),
    labelOptions = labelOptions(
      textsize  = "12px",
      direction = "auto"
    )
  )

# ── 2. Buffer each route by 25 metres ───────────────────────────────────────

# Project to a metres-based CRS first (Maryland State Plane, good for DC)
pbl_buf <- pbl_union |>
  st_transform(6487) |>          # EPSG:6487 — NAD83 / Maryland (metres)
  st_buffer(10) |>               # X m buffer
  st_transform(st_crs(cd))       # reproject back to match cd's CRS


# ── 3. Count crashes per route, month, and year within each buffer ───────────

# Spatial join: tag each crash with whichever route buffer it falls in
# (left = FALSE drops crashes outside all buffers)
crashes_in_buf <- cd |>
  st_join(pbl_buf |> select(ROUTENAME, min_build_year, max_build_year),
          join = st_within,
          left = FALSE)

# Summarise by route × year × month
crash_counts <- crashes_in_buf |>
  st_drop_geometry() |>
  group_by(ROUTENAME, year, month) |>
  summarise(
    across(all_of(injury_vars), \(x) sum(x, na.rm = TRUE)),
    n_crashes = n(),
    min_build_year = min(min_build_year, na.rm = T),
    max_build_year = max(max_build_year, na.rm = T),
    .groups   = "drop"
  ) %>%
  ungroup() %>%
  mutate(n_injury_crashes = 
           MAJORINJURIES_BICYCLIST + MINORINJURIES_BICYCLIST + UNKNOWNINJURIES_BICYCLIST + FATAL_BICYCLIST +
           MAJORINJURIES_DRIVER + MINORINJURIES_DRIVER + UNKNOWNINJURIES_DRIVER + FATAL_DRIVER +
           MAJORINJURIES_PEDESTRIAN + MINORINJURIES_PEDESTRIAN + UNKNOWNINJURIES_PEDESTRIAN + FATAL_PEDESTRIAN
           ) %>%
  mutate(n_bike_injury_crashes = 
           MAJORINJURIES_BICYCLIST + MINORINJURIES_BICYCLIST + UNKNOWNINJURIES_BICYCLIST + FATAL_BICYCLIST
           ) %>%
  mutate(n_ped_injury_crashes = 
           MAJORINJURIES_PEDESTRIAN + MINORINJURIES_PEDESTRIAN + UNKNOWNINJURIES_PEDESTRIAN + FATAL_PEDESTRIAN
           ) %>%
  select(ROUTENAME, year, month, 
         n_crashes, n_injury_crashes, n_ped_injury_crashes, n_bike_injury_crashes,
         min_build_year, max_build_year)
```


```{r}


make_dotplot <- function(custom_var=NA) {
  cc <- 
  crash_counts %>%
  filter(year %in% c(2014, 2015, 2016, 2017, 2018, 2024, 2025)) %>%
  filter(max_build_year <= 2023) %>%
  mutate(post_install = if_else(year > 2023, 1, 0)) %>%
  group_by(ROUTENAME, year) %>%
  summarise(n_crashes    = sum(!!sym(custom_var), na.rm=T),
            post_install = max(post_install)) %>%
  ungroup() %>%
  group_by(ROUTENAME, post_install) %>%
  summarise(mean_crashes = mean(n_crashes)) %>%
  ungroup()

  cc_wide <- tidyr::pivot_wider(data = cc, id_cols = "ROUTENAME", names_from = "post_install", values_from="mean_crashes")
  
  cc_wide$change <- cc_wide$`1` - cc_wide$`0`
  cc_wide <- cc_wide[!is.na(cc_wide$change),]

# ggplot(cc) +
#   geom_point(aes(x=mean_crashes, y=ROUTENAME, color=as.factor(post_install)))

ggplot(cc_wide) +
  geom_point(aes(x=change, y=ROUTENAME)) +
  geom_vline(aes(xintercept=0), color="grey") +
  geom_vline(aes(xintercept=mean(change)), color="skyblue") +
  geom_vline(aes(xintercept=median(change)), color="green") +
  labs(x="Change in the number of crashes per year") +
  ggtitle(custom_var)
}


make_dotplot("n_crashes")
make_dotplot("n_injury_crashes")
make_dotplot("n_bike_injury_crashes")
make_dotplot("n_ped_injury_crashes")

```


```{r}
# Reproject both to WGS84 for leaflet
pbl_buf_4326    <- st_transform(pbl_buf, 4326)
crashes_in_4326 <- st_transform(crashes_in_buf, 4326)

# Extract crash coordinates for circle markers
crash_coords <- st_coordinates(crashes_in_4326)

leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>

  # Route buffers
  addPolygons(
    data        = pbl_buf_4326,
    fillColor   = "steelblue",
    fillOpacity = 0.2,
    color       = "steelblue",
    weight      = 1,
    label       = ~paste0(ROUTENAME, " (built: ", min_build_year, ")"),
    labelOptions = labelOptions(textsize = "12px", direction = "auto")
  ) |>

  # Crashes
  addCircleMarkers(
    data         = crashes_in_4326,
    lng          = crash_coords[, "X"],
    lat          = crash_coords[, "Y"],
    radius       = 4,
    color        = "firebrick",
    fillColor    = "firebrick",
    fillOpacity  = 0.6,
    stroke       = FALSE,
    label        = ~paste0(ROUTENAME, " | ", month, "/", year),
    labelOptions = labelOptions(textsize = "11px", direction = "auto")
  ) |>

  addScaleBar(position = "bottomleft") |>
  addLegend(
    position = "bottomright",
    colors   = c("steelblue", "firebrick"),
    labels   = c("Route buffer (10m)", "Crash"),
    opacity  = 0.8
  )
```

Plan:

* get 10 meter buffer around every individual road segment
* count the number of crashes in each buffered segment in each year
  * if crashes fall into multiple buffers, assign them to a single buffer randomly
* merge on which segments had PBLs installed

* run regression that compares crash numbers per year, controlling for area, pre-install crashes, road width, speed limit, etc.
* compare simple before and after means on PBL & non-PBL segments



```{r}



# load roadway sub-block data, from here:
# https://opendata.dc.gov/datasets/df571ab7fea446e396bf2862d0ab6833_162/explore?location=38.927746%2C-77.024575%2C18
rs <- 
  st_read("Roadway_SubBlock.geojson") %>%
  st_zm(drop = TRUE, what = "ZM") %>%  # drops Z (and M if present) → pure XY
  # Leaflet requires WGS84 (EPSG:4326) — reproject if needed
  st_transform(4326) %>%
  st_cast("MULTILINESTRING") 

rs_buf <-
  rs |>
  st_transform(6487) |>          # EPSG:6487 — NAD83 / Maryland (metres)
  st_buffer(10) |>               # X m buffer
  st_transform(st_crs(cd))       # reproject back to match cd's CRS

rs_buf <- select(rs_buf, 
                 ROUTEID, ROUTENAME, BLOCKKEY, SUBBLOCKKEY, 
                 TOTALTRAVELLANES, TOTALPARKINGLANES, TOTALTRAVELLANEWIDTH,
                 SPEEDLIMITS_IB, SPEEDLIMITS_IB_ALT, 
                 SPEEDLIMITS_OB, SPEEDLIMITS_OB_ALT,
                 WARD_ID
                 ) %>%
  mutate(speed_limit = pmin(SPEEDLIMITS_IB, SPEEDLIMITS_IB_ALT, SPEEDLIMITS_OB, SPEEDLIMITS_OB_ALT, na.rm=T)) %>%
  select(-starts_with("SPEEDLIMITS")) %>%
  mutate(area_m = st_area(.))

# Join crashes to buffers — one crash can match multiple buffers
cd_joined <- cd |> st_join(rs_buf, join = st_within, left = FALSE)

# For crashes in multiple buffers, randomly keep one
cd_assigned <- cd_joined |>
  st_drop_geometry() |>
  group_by(CRIMEID) |>           # replace crash_id with whatever your unique crash ID column is
  slice_sample(n = 1) |>         # randomly pick one buffer per crash
  ungroup()

cd_pbl <- dplyr::left_join(cd_assigned, pbl %>% st_drop_geometry() %>% select(SUBBLOCKKE, ROUTENAME, build_year), by=c("SUBBLOCKKEY"="SUBBLOCKKE"))

units(cd_pbl$area_m) <- NULL

cd_pbl_clean <-
  cd_pbl %>%
  st_drop_geometry() %>%
  group_by(SUBBLOCKKEY, year) %>%
  summarise(
    across(all_of(injury_vars), \(x) sum(x, na.rm = TRUE)),
    n_crashes            = n(),
    min_build_year       = min(build_year, na.rm = T),
    max_build_year       = max(build_year, na.rm = T),
    TOTALTRAVELLANES     = median(TOTALTRAVELLANES, na.rm=T),
    TOTALPARKINGLANES    = median(TOTALPARKINGLANES, na.rm=T),
    TOTALTRAVELLANEWIDTH = median(TOTALTRAVELLANEWIDTH, na.rm=T),
    WARD_ID              = median(WARD_ID, na.rm=T),
    speed_limit          = median(speed_limit, na.rm=T),
    area_m               = median(area_m, na.rm=T),
    .groups   = "drop"
  ) %>%
  ungroup() %>%
  mutate(
    crashes_per_acre     = n_crashes / area_m * 4046.86
  )
```


































