---
title: "DDOT project evaluations"
author: "pete rodrigue"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---


DDOT conducts evaluations of their bus priority and bike infrastruture projects. Here's a summary of what those evaluations have found.

```{r setup, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, message=F, warning=F)

setwd("~/GitHub/ddot-project-evals")

# Libraries we need:
library(readr)
library(ggplot2)
library(dplyr)
library(stringr)
library(purrr)
library(broom)
library(sf)
library(leaflet)
library(tidyr)
library(tidyverse)

# google sheet w/ the data:
url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vQtDsTt9VVyn-wptsMT6vxtCM2J1dPYWji90OOWj1v4c4pPz6Kwa6mQtOk096EU3npvJ9g4fcjnwLpp/pub?gid=1445637369&single=true&output=csv"

# load the data from the google sheet:
df <- read_csv(url, show_col_types = F)
```


```{r}
to_plot <-
  df %>%
  filter(Outcome %in% c("Injury crashes", "All crashes")) %>%
  mutate(Change = Change*-1)

ggplot(to_plot) +
  geom_bar(aes(x=`Project name`, y=Change, fill=Outcome), stat="identity", position="dodge") +
  labs(y="% decrease in crashes", x="") +
  ggtitle("All projects saw decreases in crashes") +
  geom_hline(aes(yintercept=mean(Change[Outcome == "All crashes"], na.rm=T)), color="#F8766D", linetype="dashed", linewidth=.8, alpha=.5) +
  geom_hline(aes(yintercept=mean(Change[Outcome == "Injury crashes"], na.rm=T)), color="#00BFC4", linetype="dashed", linewidth=.8, alpha=.5) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

```{r}
to_plot <-
  df %>%
  filter(Outcome %in% c("Pedestrian injury crashes")) %>%
  mutate(Change = Change*-1)

ggplot(to_plot) +
  geom_bar(aes(x=`Project name`, y=Change), stat="identity", fill="seagreen") +
  labs(y="% decrease in crashes", x="") +
  ggtitle("Pedestrian crashes fell after project completion, sometimes to zero") +
  geom_hline(aes(yintercept=mean(Change, na.rm=T)), color="grey", linetype="dashed", size=1) +
  theme_minimal() +
  theme(legend.title = element_blank())
```



```{r}
to_plot <-
  df %>%
  filter(str_detect(Outcome, "Private motor vehicle")) %>%
  group_by(`Project name`, Outcome) %>%
  summarize(change = mean(Change, na.rm=T)) %>%
  ungroup() %>%
  mutate(Outcome = ifelse(Outcome == "Private motor vehicle AM peak travel time (seconds)", 
                          "AM traffic travel time",
                          "PM traffic travel time"))

ggplot(to_plot) +
  geom_point(
    aes(x=change, color=Outcome, y=`Project name`), size=3, alpha=.7
  ) +
  geom_vline(aes(xintercept=0), color="black") +
  geom_vline(aes(xintercept=mean(change[Outcome == "AM traffic travel time"], na.rm=T)), color="#F8766D", linetype="dashed") +
  geom_vline(aes(xintercept=mean(change[Outcome == "PM traffic travel time"], na.rm=T)), color="#00BFC4", linetype="dashed") +
  labs(x="Change in travel time in seconds", y="") +
  theme_minimal() +
  xlim(c(min(to_plot$change)-10, max(to_plot$change)+25)) +
  theme(legend.title = element_blank()) +
  ggtitle("Drivers' travel times faster on average after project completion")
```


```{r}
rm(df); rm(to_plot); rm(url)
```

<br><br>

DDOT looked at a few projects in depth. But we can also use DC's crash data to look at every protected and buffered bike lane in DDOT's "bike lane" shapefile.

```{r load_clean_crash_data}
# load crash data
# data downloaded from here:
# https://opendata.dc.gov/datasets/crashes-in-dc/about
cd <- readr::read_csv("data/Crashes_in_DC.csv", show_col_types = F)

# create year and month of crash variables:
cd$year  <- as.numeric(substr(x = cd$REPORTDATE, start = 1, stop = 4))
cd$month <- as.numeric(substr(x = cd$REPORTDATE, start = 6, stop = 7))

# remove rows with no date information:
cd <- filter(cd, !is.na(year))

# this data has crashes going back to the 2000s, but the data doesn't look super complete until 2014 or so:
# table(cd$year)
# According to DDOT, the location information in the data set became much more accurate starting in 2016:
# https://ddotwiki.atlassian.net/wiki/spaces/GIS0225/pages/2053603429/Crash+Data

# project crash data:
cd <- cd |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

# injury variables we care about:
injury_vars <- c(
  "MAJORINJURIES_BICYCLIST",   "MINORINJURIES_BICYCLIST",
  "UNKNOWNINJURIES_BICYCLIST", "FATAL_BICYCLIST",
  "MAJORINJURIES_DRIVER",      "MINORINJURIES_DRIVER",
  "UNKNOWNINJURIES_DRIVER",    "FATAL_DRIVER",
  "MAJORINJURIES_PEDESTRIAN",  "MINORINJURIES_PEDESTRIAN",
  "UNKNOWNINJURIES_PEDESTRIAN","FATAL_PEDESTRIAN",
  "TOTAL_VEHICLES", "TOTAL_BICYCLES", "TOTAL_PEDESTRIANS"
)

# subset to just variables we care about:
cd <- select(cd, CRIMEID, year, month, injury_vars)

coords <- st_coordinates(cd)
cd$crash_lon <- coords[, "X"]
cd$crash_lat <- coords[, "Y"]
```


```{r load_clean_bl_data}
# load labeled PBL data that has the road subblock IDs and the year in which 
# PBLs were installed on those subblocks (we hand labeled them)
pbl <- 
  st_read("data/Protected-Buffered-Bike-Lanes.geojson", quiet=T) %>%
  filter(!is.na(install_year)) %>% 
  filter(install_year != 99) %>%
  rename(build_year = install_year) #%>%
  # filter((!is.na(BIKELANE_PROTECTED)) | (!is.na(BIKELANE_DUAL_PROTECTED)))
```




```{r load_clean_cabi_data}
# go to this other file to download and clean the Cabi data if you need to:
# source("clean_cabi_data.R")

# otherwise just load the cleaned data:
cabi <- readr::read_csv("data/cabi_both_years_clean.csv", show_col_types = F)

cabi <-
  cabi |>
  mutate(
    n_trips_ended_2018 = if_else(is.na(n_trips_ended_2018), 0, n_trips_ended_2018),
    n_trips_ended_2025 = if_else(is.na(n_trips_ended_2025), 0, n_trips_ended_2025),
    change_n_trips     = n_trips_ended_2025 - n_trips_ended_2018
  ) %>%
  filter(!is.na(end_lat))

# # Diverging palette centered at 0 — red = fewer trips, blue = more trips
# pal <- colorNumeric(
#   palette = "RdBu",
#   domain  = cabi$change_n_trips,
#   reverse = FALSE
# )
# leaflet(cabi) |>
#   addProviderTiles(providers$CartoDB.Positron) |>
#   addCircleMarkers(
#     lng          = ~end_lng,
#     lat          = ~end_lat,
#     radius       = 6,
#     color        = "white",
#     weight       = 0.5,
#     fillColor    = ~pal(change_n_trips),
#     fillOpacity  = 0.85,
#     label = ~paste0(
#       "<b>", `End station number`, "</b><br>",
#       "2018 trips: ", scales::comma(n_trips_ended_2018), "<br>",
#       "2025 trips: ", scales::comma(n_trips_ended_2025), "<br>",
#       "Change: ", scales::comma(change_n_trips)
#     ) |> lapply(htmltools::HTML),
#     labelOptions = labelOptions(textsize = "12px", direction = "auto")
#   ) |>
#   addLegend(
#     position = "bottomright",
#     pal      = pal,
#     values   = ~change_n_trips,
#     title    = "Change in<br>annual trips",
#     labFormat = labelFormat(big.mark = ","),
#     opacity  = 0.9
#   ) |>
#   addScaleBar(position = "bottomleft")

# Convert cabi to sf object
cabi_sf <- cabi |>
  filter(!is.na(end_lat), !is.na(end_lng)) |>
  st_as_sf(coords = c("end_lng", "end_lat"), crs = 4326) |>
  st_transform(st_crs(4326))


# add rough CaBi ride totals by year:
annual_cabi_rides <-
  data.frame(
    "year"       = c(2017, 2024, 2025),
    "cabi_rides" = c(3.8 , 6.1 , 6.6)
    )
```

what we'll do:

* get 10 meter buffer around every individual road segment
* count the number of crashes in each buffered segment in each year
  * if crashes fall into multiple buffers, assign them to a single buffer randomly. this adds noise and will attenuate the "treatment effect"
* merge on which segments had PBLs & buffered bike lanes installed

* run regression that compares crash numbers per year, controlling for area, pre-install crashes, road width, bike traffic from Cabi system, etc
* compare simple before and after means on PBL/BBL & non-PBL/BBL segments


```{r load_clean_subblock_data}

# load all roadway sub-block data, from here:
# https://opendata.dc.gov/datasets/df571ab7fea446e396bf2862d0ab6833_162/explore?location=38.927746%2C-77.024575%2C18
# This will be our main data set we merge other data onto.
rs <- 
  st_read("data/Roadway_SubBlock.geojson", quiet=T) %>%
  filter(ROADTYPE == 1) %>%             # keep just streets, remove service roads, ramps, alleys, driveways, trails, and walkways
  st_zm(drop = TRUE, what = "ZM") %>%   # drops Z (and M if present) → pure XY
  st_transform(4326) %>%                # Leaflet requires WGS84 (EPSG:4326) — reproject if needed
  st_cast("MULTILINESTRING")
```

```{r merge_subblock_cabi_data}
# create an X meter buffer around each subblock to count the number of nearby 
# cabi trips
rs_cabi_buf <-
  rs |>
  st_transform(6487) |>          # EPSG:6487 — NAD83 / Maryland (metres)
  st_buffer(dist = 500) |>              # X m buffer
  st_transform(st_crs(cd))       # reproject back to match cd's CRS

# Spatial join stations to buffers, then summarise by subblock
rs_cabi_change <- rs_cabi_buf |>
  st_join(cabi_sf, join = st_contains) |>
  st_drop_geometry() |>
  group_by(SUBBLOCKKEY) |>
  summarise(
    total_change_trips = sum(change_n_trips, na.rm = TRUE),
    n_stations         = sum(!is.na(change_n_trips)),
    .groups = "drop"
  )
```


```{r merge_subblock_crash_data}
# OK now let's join the crashes to the subblocks, which we'll do with anther,
# smaller, X meter buffer:
rs_radius <- (rs$TOTALPARKINGLANES + rs$TOTALTRAVELLANES)*4 / 2 + 4
rs_crash_buf <-
  rs |>
  st_transform(6487) |>                                       # EPSG:6487 — NAD83 / Maryland (meters)
  st_buffer(dist = rs_radius, endCapStyle = "FLAT") |>        # X m buffer
  st_transform(st_crs(cd)) |>                                 # reproject back to match cd's CRS
  select(ROUTEID, ROUTENAME, BLOCKKEY, SUBBLOCKKEY, 
         TOTALTRAVELLANES, TOTALPARKINGLANES, TOTALTRAVELLANEWIDTH,
         WARD_ID, DCFUNCTIONALCLASS
         ) %>%
  mutate(area_m = st_area(.))

units(rs_crash_buf$area_m) <- NULL

leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    data = rs_crash_buf %>% 
            st_crop(xmin = -77.03, ymin = 38.94, xmax = -77.01, ymax = 38.944), 
    fillColor = "blue", 
    stroke=F) |>
  addCircleMarkers(
    data = cd %>% 
      st_crop(xmin = -77.03, ymin = 38.94, xmax = -77.010, ymax = 38.944) %>%
      filter(year==2025),
    lat=~crash_lat, 
    lng=~crash_lon, 
    stroke=F, color="red", radius = 4, label=)
```


```{r}
# Join crashes to buffers — one crash can match multiple buffers, since the buffers can overlap
cd_joined <- cd |> st_join(rs_crash_buf, join = st_within, left = FALSE)

# For crashes in multiple buffers, randomly keep one. this will add noise but that's probably OK.
cd_assigned <- cd_joined |>
  st_drop_geometry() #|>
  # group_by(CRIMEID) |>           
  # slice_sample(n = 1) |>         # randomly pick one buffer per crash
  # ungroup()


# count up the number of crashes in each subblock
cd_assigned_clean <-
  cd_assigned %>%
  mutate(WARD_ID = as.numeric(WARD_ID)) %>%
  group_by(SUBBLOCKKEY, year) %>%
  summarise(
    across(all_of(injury_vars), \(x) sum(x, na.rm = TRUE)),
    n_crashes            = n(),
    TOTALTRAVELLANES     = first(TOTALTRAVELLANES),
    TOTALPARKINGLANES    = first(TOTALPARKINGLANES),
    TOTALTRAVELLANEWIDTH = first(TOTALTRAVELLANEWIDTH),
    WARD_ID              = first(WARD_ID),
    DCFUNCTIONALCLASS    = first(DCFUNCTIONALCLASS),
    area_m               = first(area_m),
    .groups   = "drop"
  ) %>%
  ungroup() %>%
  mutate(
    crashes_per_acre     = n_crashes / area_m * 4046.86
  ) %>%
  mutate(across(where(is.numeric), ~ ifelse(is.nan(.), NA, .)))

# subset to real years:
cd_assigned_clean <- filter(cd_assigned_clean, (year > 2015) & (year <= 2025))

# ensures the data frame is a balanced panel
# i.e., that every subblock has a row for every year from 2015 to 2025, with no gaps.
cd_assigned_clean <- cd_assigned_clean |>
  complete(
    SUBBLOCKKEY,
    year = 2015:2025,
    fill = list(
      n_crashes            = 0,
      TOTALTRAVELLANES     = NA_real_,
      TOTALPARKINGLANES    = NA_real_,
      TOTALTRAVELLANEWIDTH = NA_real_,
      WARD_ID              = NA_real_,
      area_m               = NA_real_,
      DCFUNCTIONALCLASS    = NA_real_,
      crashes_per_acre     = 0,
      # injury vars
      MAJORINJURIES_BICYCLIST   = 0,
      MINORINJURIES_BICYCLIST   = 0,
      UNKNOWNINJURIES_BICYCLIST = 0,
      FATAL_BICYCLIST           = 0,
      MAJORINJURIES_DRIVER      = 0,
      MINORINJURIES_DRIVER      = 0,
      UNKNOWNINJURIES_DRIVER    = 0,
      FATAL_DRIVER              = 0,
      MAJORINJURIES_PEDESTRIAN  = 0,
      MINORINJURIES_PEDESTRIAN  = 0,
      UNKNOWNINJURIES_PEDESTRIAN= 0,
      FATAL_PEDESTRIAN          = 0,
      TOTAL_VEHICLES            = 0,
      TOTAL_BICYCLES            = 0,
      TOTAL_PEDESTRIANS         = 0
    )
  )

# fill the time invariant data into the gaps:
cd_assigned_clean <- cd_assigned_clean |>
  arrange(SUBBLOCKKEY, year) |>
  group_by(SUBBLOCKKEY) |>
  fill(TOTALTRAVELLANES, TOTALPARKINGLANES, TOTALTRAVELLANEWIDTH, DCFUNCTIONALCLASS,
       WARD_ID, area_m, .direction = "downup") |>
  ungroup()

# (finally!) join on the protected and buffered bike lane info:
cd_pbl_clean <- 
  dplyr::left_join(cd_assigned_clean,
                   pbl %>% 
                     st_drop_geometry() %>%
                     select(SUBBLOCKKEY, build_year),
                   by="SUBBLOCKKEY"
                   )

# clean up our data, remove COVID years. 
# Our last 2 years (2024 and 2025) will serve as our post-treatment years for most subblocks.
cd_pbl_clean <- cd_pbl_clean %>%
  mutate(build_year = if_else(build_year == 99, NA, build_year)) %>%
  filter(year %in% c(2016, 2017, 2018, 2019, 2024, 2025)) %>%
  mutate(post_install = if_else((year > 2023) | (year >= build_year), 1, 0)) 

cd_pbl_clean$post_install[(cd_pbl_clean$year >= 2024) & is.na(cd_pbl_clean$build_year)] <- 1
cd_pbl_clean$post_install[(cd_pbl_clean$year  < 2024) & is.na(cd_pbl_clean$build_year)] <- 0

# filter out crash counts that are probably data errors.
# it seems unlikely that this one stretch of road in ward 7 had over 250 crashes in a year
cd_pbl_clean <- filter(cd_pbl_clean, n_crashes < 250)

cd_pbl_clean <-
  cd_pbl_clean %>%
  group_by(SUBBLOCKKEY) %>%
  mutate(
    min_post_val = min(post_install, na.rm=T),
    max_post_val = max(post_install, na.rm=T)
         ) %>%
  ungroup() %>%
    filter((min_post_val==0) & (max_post_val==1)) %>%
  mutate(
    n_injury_crashes = 
           MAJORINJURIES_BICYCLIST + MINORINJURIES_BICYCLIST + UNKNOWNINJURIES_BICYCLIST + FATAL_BICYCLIST +
           MAJORINJURIES_DRIVER + MINORINJURIES_DRIVER + UNKNOWNINJURIES_DRIVER + FATAL_DRIVER +
           MAJORINJURIES_PEDESTRIAN + MINORINJURIES_PEDESTRIAN + UNKNOWNINJURIES_PEDESTRIAN + FATAL_PEDESTRIAN,
    n_driver_injury_crashes = 
           MAJORINJURIES_DRIVER + MINORINJURIES_DRIVER + UNKNOWNINJURIES_DRIVER + FATAL_DRIVER,
    n_serious_driver_injury_crashes = 
           MAJORINJURIES_DRIVER + FATAL_DRIVER,
    n_bike_injury_crashes = 
           MAJORINJURIES_BICYCLIST + MINORINJURIES_BICYCLIST + UNKNOWNINJURIES_BICYCLIST + FATAL_BICYCLIST,
    n_serious_bike_injury_crashes = 
           MAJORINJURIES_BICYCLIST + FATAL_BICYCLIST,
    n_ped_injury_crashes = 
           MAJORINJURIES_PEDESTRIAN + MINORINJURIES_PEDESTRIAN + UNKNOWNINJURIES_PEDESTRIAN + FATAL_PEDESTRIAN,
    n_serious_ped_injury_crashes = 
           MAJORINJURIES_PEDESTRIAN + FATAL_PEDESTRIAN,
    
    injury_crashes_per_acre                = n_injury_crashes / area_m * 4046.86,
    driver_injury_crashes_per_acre         = n_driver_injury_crashes / area_m * 4046.86,
    serious_driver_injury_crashes_per_acre = n_serious_driver_injury_crashes / area_m * 4046.86,
    bike_injury_crashes_per_acre           = n_bike_injury_crashes / area_m * 4046.86,
    serious_bike_injury_crashes_per_acre   = n_serious_bike_injury_crashes / area_m * 4046.86,
    ped_injury_crashes_per_acre            = n_ped_injury_crashes / area_m  * 4046.86,
    serious_ped_injury_crashes_per_acre    = n_serious_ped_injury_crashes / area_m  * 4046.86,
    
    total_vehicles_per_acre                = TOTAL_VEHICLES / area_m  * 4046.86,
    total_bikes_per_acre                   = TOTAL_BICYCLES / area_m  * 4046.86,
    total_peds_per_acre                    = TOTAL_PEDESTRIANS / area_m  * 4046.86,
    
    treatment = if_else(is.na(build_year), 0, 1)
           ) %>%
  group_by(SUBBLOCKKEY) %>%
  mutate(
    mean_crashes_pre2019 = mean(crashes_per_acre[year < 2019], na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(mean_crashes_pre2019 = na_if(mean_crashes_pre2019, NaN)) %>%
  group_by(SUBBLOCKKEY) %>%
  mutate(mean_crashes_pre2019 = mean(mean_crashes_pre2019, na.rm=T)) %>%
  ungroup()



cd_pbl_clean <- dplyr::left_join(cd_pbl_clean, rs_cabi_change, by="SUBBLOCKKEY")


cd_pbl_clean <- dplyr::left_join(cd_pbl_clean, annual_cabi_rides, by="year")

run_regs <- function(df, outcome, include_year=T, return_mod=F) {
  print(paste("############### ", outcome))
  print("################################")
  root_reg <- paste0(outcome, " ~ treatment + post_install + treatment*post_install")
  print(summary(lm(formula = as.formula(root_reg), data = df)))
  cat("\n****************************************************")
  print(summary(lm(formula = as.formula(paste0(root_reg, " + mean_crashes_pre2019")), data = df)))
  cat("\n****************************************************")
  print(summary(lm(formula = as.formula(paste0(root_reg, " + mean_crashes_pre2019", 
        ifelse(include_year, " + as.factor(year)", ""))), data = df)))
  cat("\n****************************************************")
  full_mod <- lm(formula = as.formula(
    paste0(root_reg, " + mean_crashes_pre2019 + TOTALTRAVELLANEWIDTH + TOTALPARKINGLANES ",
          "+ as.factor(WARD_ID) + as.factor(DCFUNCTIONALCLASS) + total_change_trips + cabi_rides",
           ifelse(include_year, " + as.factor(year)", ""))),
    data = df)
  print(summary(full_mod))

  if (return_mod) {
    return(full_mod)
  }
}

# run_regs(df = cd_pbl_clean, outcome = "crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "driver_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_driver_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "bike_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_bike_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "ped_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_ped_injury_crashes_per_acre")
# 
# to_plot <- run_regs(df = cd_pbl_clean, outcome = "injury_crashes_per_acre", return_mod=T)
# plot(to_plot)

cd_pbl_clean_simplified <-
  cd_pbl_clean %>%
  group_by(SUBBLOCKKEY, post_install) %>%
  summarize(
    TOTALTRAVELLANEWIDTH = mean(TOTALTRAVELLANEWIDTH, na.rm=T),
    TOTALPARKINGLANES    = mean(TOTALPARKINGLANES, na.rm=T),
    DCFUNCTIONALCLASS    = mean(DCFUNCTIONALCLASS, na.rm=T),
    WARD_ID              = mean(WARD_ID, na.rm=T),
    total_change_trips   = mean(total_change_trips, na.rm=T),
    
    across(ends_with("_per_acre"), \(x) mean(x, na.rm = TRUE)),
    
    treatment                            = mean(treatment, na.rm=T),
    mean_crashes_pre2019                 = mean(mean_crashes_pre2019, na.rm=T)
            ) %>%
  ungroup()

# run_regs(df = cd_pbl_clean_simplified, outcome = "crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "driver_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "serious_driver_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "bike_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "serious_bike_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "ped_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "serious_ped_injury_crashes_per_acre", F)
```


```{r, fig.width=12}
run_regs <- function(df, outcome, include_year = TRUE) {
  root_reg <- paste0(outcome, " ~ treatment + post_install + treatment*post_install")
  year_term <- ifelse(include_year, " + as.factor(year)", "")

  formulas <- list(
    simple = as.formula(root_reg),
    full   = as.formula(paste0(
      root_reg,
      " + mean_crashes_pre2019 + TOTALTRAVELLANEWIDTH + TOTALPARKINGLANES",
      " + as.factor(WARD_ID) + as.factor(DCFUNCTIONALCLASS) + total_change_trips",
      year_term
    ))
  )

  map_dfr(formulas, \(f) {
    tidy(lm(f, data = df)) |>
      filter(term == "treatment:post_install") |>
      mutate(outcome = outcome)
  }, .id = "model")
}


plot_effects <- function(outcomes) {
  results <- map_dfr(outcomes, \(o) {
    run_regs(df = cd_pbl_clean_simplified, outcome = o, include_year = FALSE)
  })
  
  results <- results |>
    mutate(
      significance = case_when(
        p.value < 0.001 ~ "p < 0.001",
        p.value < 0.01  ~ "p < 0.01",
        p.value < 0.05  ~ "p < 0.05",
        TRUE            ~ "p ≥ 0.05"
      ),
      significance = factor(significance,
                            levels = c("p < 0.001", "p < 0.01",
                                       "p < 0.05",  "p ≥ 0.05")),
      model  = factor(model, levels = c("simple", "full"),
                      labels = c("Simple model", "Full model")),
      outcome = factor(outcome, levels = outcomes) |> # preserve ordering
        fct_relabel(\(x) str_replace_all(x, "_", " ") |>
                                str_replace(" per acre", "") |>
                                str_to_sentence())
    )
  
  
  sig_colors <- c(
    "p < 0.001" = "#2166ac",
    "p < 0.01"  = "#74add1",
    "p < 0.05"  = "#fdae61",
    "p ≥ 0.05"  = "#d9d9d9"
  )
  
  ggplot(
         filter(results, model=="Simple model"), 
         aes(x = outcome, y = estimate, fill = significance)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_errorbar(
      aes(ymin = estimate - 1.96 * std.error,
          ymax = estimate + 1.96 * std.error),
      position = position_dodge(width = 0.7),
      width = 0.2, color = "grey30"
    ) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
    scale_fill_manual(values = sig_colors, drop = FALSE) +
    scale_x_discrete(labels = \(x) str_replace_all(x, "_", "\n")) +
    # facet_wrap(~model) +
    labs(
      title = "Effect of protected bike lane installation on crash outcomes",
      subtitle = "Coefficient on treatment × post-install interaction",
      x     = NULL,
      y     = "Estimated effect per acre",
      fill  = "Significance"
    ) +
    theme_minimal() +
    theme(
      axis.text.x     = element_text(size = 8),
      legend.position = "bottom",
      panel.grid.major.x = element_blank()
    )
}


plot_effects(
  outcomes = c(
      "crashes_per_acre",
      "injury_crashes_per_acre",
      
      "driver_injury_crashes_per_acre",
      "serious_driver_injury_crashes_per_acre",
  
      "bike_injury_crashes_per_acre",
      "serious_bike_injury_crashes_per_acre",
      
      "ped_injury_crashes_per_acre",
      "serious_ped_injury_crashes_per_acre"
      )
  )

plot_effects(
  outcomes = c(
      "total_vehicles_per_acre",
      "total_bikes_per_acre",
      "total_peds_per_acre"
      )
  )
```



```{r}
  outcomes = c(
      "crashes_per_acre",
      "injury_crashes_per_acre",
      
      "driver_injury_crashes_per_acre",
      "serious_driver_injury_crashes_per_acre",
  
      "bike_injury_crashes_per_acre",
      "serious_bike_injury_crashes_per_acre",
      
      "ped_injury_crashes_per_acre",
      "serious_ped_injury_crashes_per_acre"
      )

summary_df <-
  cd_pbl_clean_simplified |>
  select(SUBBLOCKKEY, post_install, treatment, ends_with("_per_acre")) |>
  pivot_longer(cols = all_of(outcomes),
               names_to  = "outcome",
               values_to = "value") |>
  group_by(outcome, post_install, treatment) |>        # add treatment to grouping
  summarise(
    mean = mean(value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    post_install = factor(post_install, levels = c(0, 1),
                          labels = c("Pre", "Post")),
    treatment = factor(treatment, levels = c(0, 1),
                       labels = c("Other streets", "PBL installs")),
    outcome = factor(outcome, levels = outcomes) |>
                fct_relabel(\(x) str_replace_all(x, "_", " ") |>
                              str_replace(" per acre", "") |>
                              str_to_sentence())
  )

ggplot(summary_df, aes(x = post_install, y = mean,
                       color = treatment, group = treatment)) +
  geom_point(size = 3) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~outcome, nrow = 1,
             labeller = labeller(outcome = label_wrap_gen(15))) +
  scale_color_manual(values = c("Other streets"   = "grey50",
                                "PBL installs" = "#4575b4")) +
  labs(
    title = "Mean annual crashes per acre before and after bike lane installation",
    x     = NULL,
    y     = "Crashes per acre",
    color = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position    = "bottom",
    panel.grid.major.x = element_blank(),
    strip.text         = element_text(size = 7)
  )
```




```{r}

# # Get unique treated subblock keys
# treated_keys <- cd_pbl_clean_simplified |>
#   filter(treatment == 1) |>
#   distinct(SUBBLOCKKEY)
# 
# # Merge onto rs geometry
# rs_treatment <- rs |>
#   inner_join(treated_keys, by = "SUBBLOCKKEY")
# 
# # Plot
# leaflet(rs_treatment |> st_zm(drop = TRUE)) |>
#   addProviderTiles(providers$CartoDB.Positron) |>
#   addPolylines(
#     color   = "#4575b4",
#     weight  = 2.5,
#     opacity = 0.85,
#     label   = ~SUBBLOCKKEY,
#     labelOptions = labelOptions(textsize = "12px", direction = "auto"),
#     highlightOptions = highlightOptions(
#       weight      = 7,
#       color       = "#222222",
#       opacity     = 1,
#       bringToFront = TRUE
#     )
#   )

```

























