---
title: "DDOT project evaluations"
author: "pete rodrigue"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---


DDOT conducts evaluations of their bus priority and bike infrastruture projects. Here's a summary of what those evaluations have found.

```{r setup, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, message=F, warning=F)

setwd("~/GitHub/ddot-project-evals")

# Libraries we need:
library(readr)
library(ggplot2)
library(dplyr)
library(stringr)
library(purrr)
library(broom)
library(sf)
library(leaflet)
library(tidyr)
library(tidyverse)
library(plotly)
library(zoo)

# google sheet w/ the data:
url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vQtDsTt9VVyn-wptsMT6vxtCM2J1dPYWji90OOWj1v4c4pPz6Kwa6mQtOk096EU3npvJ9g4fcjnwLpp/pub?gid=1445637369&single=true&output=csv"

# load the data from the google sheet:
df <- read_csv(url, show_col_types = F)
```


```{r}
to_plot <-
  df %>%
  filter(Outcome %in% c("Injury crashes", "All crashes")) %>%
  mutate(Change = Change*-1)

ggplot(to_plot) +
  geom_bar(aes(x=`Project name`, y=Change, fill=Outcome), stat="identity", position="dodge") +
  labs(y="% decrease in crashes", x="") +
  ggtitle("All projects saw decreases in crashes") +
  geom_hline(aes(yintercept=mean(Change[Outcome == "All crashes"], na.rm=T)), 
             color="#F8766D", linetype="dashed", linewidth=.8, alpha=.5) +
  geom_hline(aes(yintercept=mean(Change[Outcome == "Injury crashes"], na.rm=T)), 
             color="#00BFC4", linetype="dashed", linewidth=.8, alpha=.5) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

```{r}
to_plot <-
  df %>%
  filter(Outcome %in% c("Pedestrian injury crashes")) %>%
  mutate(Change = Change*-1)

ggplot(to_plot) +
  geom_bar(aes(x=`Project name`, y=Change), stat="identity", fill="seagreen") +
  labs(y="% decrease in crashes", x="") +
  ggtitle("Pedestrian crashes fell after project completion, sometimes to zero") +
  geom_hline(aes(yintercept=mean(Change, na.rm=T)), color="grey", linetype="dashed", size=1) +
  theme_minimal() +
  theme(legend.title = element_blank())
```



```{r}
to_plot <-
  df %>%
  filter(str_detect(Outcome, "Private motor vehicle")) %>%
  group_by(`Project name`, Outcome) %>%
  summarize(change = mean(Change, na.rm=T)) %>%
  ungroup() %>%
  mutate(Outcome = ifelse(Outcome == "Private motor vehicle AM peak travel time (seconds)", 
                          "AM traffic travel time",
                          "PM traffic travel time"))

ggplot(to_plot) +
  geom_point(
    aes(x=change, color=Outcome, y=`Project name`), size=3, alpha=.7
  ) +
  geom_vline(aes(xintercept=0), color="black") +
  geom_vline(aes(xintercept=mean(change[Outcome == "AM traffic travel time"], na.rm=T)), color="#F8766D", linetype="dashed") +
  geom_vline(aes(xintercept=mean(change[Outcome == "PM traffic travel time"], na.rm=T)), color="#00BFC4", linetype="dashed") +
  labs(x="Change in travel time in seconds", y="") +
  theme_minimal() +
  xlim(c(min(to_plot$change)-10, max(to_plot$change)+25)) +
  theme(legend.title = element_blank()) +
  ggtitle("Drivers' travel times faster on average after project completion")
```


```{r}
rm(df); rm(to_plot); rm(url)
```


```{r}
outcome_vars <-
  c("All Crashes", "Vehicle Crashes", "Bike Crashes", "Pedestrian Crashes", 
    "Any Injury", "Driver Injury", "Bicyclist Injury", "Pedestrian Injury",
    "Driver Serious Injury", "Bicyclist Serious Injury", "Pedestrian Serious Injury")

# load data:
source("load_data.R")
# load crash data:
cd <- load_crash_data()
# crash_details <- read_csv("data/Crash_Details_Table.csv", show_col_types = F)
# load project shapefiles:
ddot_projs <- 
  st_read("data/ddot-evaluated-projects.geojson", quiet=T) |>
  st_transform(4326) 

ddot_projs_buf <-
  ddot_projs |>
  st_transform(6487) |>               # EPSG:6487 — NAD83 / Maryland (metres)
  st_buffer(dist = 4,                # X m buffer
            endCapStyle = "FLAT"
            ) |>  
  st_transform(st_crs(ddot_projs))    # reproject back to match cd's CRS

crashes_near_projects <- st_join(cd, ddot_projs_buf, join = st_within, left = FALSE)

# plot the projects:
source("make_plots.R")

plot_leaflet(
  lines_sf = ddot_projs, 
  polygons_sf = ddot_projs_buf, 
  points_sf = crashes_near_projects
  )


cd_in_projs_summed <-
  crashes_near_projects |>
  st_drop_geometry()    |>
  group_by(project_name, month, year, date) |>
  summarize(
    `All Crashes` = n(),
    across(all_of(outcome_vars), \(x) sum(x, na.rm = TRUE)),
    .groups = "drop_last"
    ) |>
  ungroup()


# Minnesota Ave dates:
# Before: Feb 2017 - Jan 2020
# After:  Nov 2023 - Oct 2024

```

```{r, fig.width=10, fig.height=8}
plot_crashes_over_time <- function(data, smoothing_factor = 1,
                                   period_1 = NULL, period_2 = NULL,
                                   title = NULL) {
  
  series_levels <- c("All Crashes", "Any Injury",
                     "Vehicle Crashes", "Bike Crashes",      "Pedestrian Crashes",
                     "Driver Injury",   "Bicyclist Injury",  "Pedestrian Injury")
  
  series_colors <- c(
    "All Crashes"        = "#636363",
    "Any Injury"         = "#bdbdbd",
    "Vehicle Crashes"    = "#2166ac",
    "Driver Injury"      = "#74add1",
    "Bike Crashes"       = "#1a7837",
    "Bicyclist Injury"   = "#7fbf7b",
    "Pedestrian Crashes" = "#d7191c",
    "Pedestrian Injury"  = "#f4a582"
  )
  
  panel_map <- c(
    "All Crashes"        = "Overall",
    "Any Injury"         = "Overall",
    "Vehicle Crashes"    = "Vehicle",
    "Driver Injury"      = "Vehicle",
    "Bike Crashes"       = "Bike",
    "Bicyclist Injury"   = "Bike",
    "Pedestrian Crashes" = "Pedestrian",
    "Pedestrian Injury"  = "Pedestrian"
  )
  
  # ── 1. Compute series ───────────────────────────────────────────────────────
  plot_data <- data |>
    select(date, all_of(series_levels)) |>
    pivot_longer(-date, names_to = "series", values_to = "n") |>
    arrange(series, date) |>
    group_by(series) |>
    mutate(value_smooth = rollmean(n, k = smoothing_factor, fill = NA, align = "right")) |>
    ungroup() |>
    mutate(
      series = factor(series, levels = series_levels),
      panel  = factor(panel_map[as.character(series)],
                      levels = c("Overall", "Vehicle", "Bike", "Pedestrian"))
    )
  
  # ── 2. Build plot ───────────────────────────────────────────────────────────
  p <- ggplot(plot_data, aes(x = date, y = value_smooth, color = series)) +
    geom_line(linewidth = 0.7, alpha = 0.8, na.rm = TRUE) +
    facet_wrap(~ panel, ncol = 1, scales = "free_y") +
    scale_color_manual(values = series_colors) +
    scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
    labs(x = NULL, y = "Count", color = NULL) +
    theme_minimal(base_size = 11) +
    theme(
      legend.position  = "bottom",
      strip.text       = element_text(face = "bold", size = 10),
      axis.text.x      = element_text(angle = 45, hjust = 1),
      panel.grid.minor = element_blank()
    ) +
    guides(color = guide_legend(nrow = 2)) +
    ggtitle(title)
  
  # ── 3. Add period shading ───────────────────────────────────────────────────
  p <- p +
    annotate("rect",
      xmin = as.Date(period_1$start), xmax = as.Date(period_1$end),
      ymin = -Inf, ymax = Inf,
      fill = "#2166ac", alpha = 0.08
    ) 

  p <- p +
    annotate("rect",
      xmin = as.Date(period_2$start), xmax = as.Date(period_2$end),
      ymin = -Inf, ymax = Inf,
      fill = "#d7191c", alpha = 0.08
    ) 
  
  # ── 4. Summary table: outcomes per month in each period ─────────────────────
  summarise_period <- function(p) {
    n_months <- interval(as.Date(p$start), as.Date(p$end)) %/% months(1) + 1
    
    plot_data |>
      filter(date >= as.Date(p$start), date <= as.Date(p$end)) |>
      group_by(series) |>
      summarise(total = sum(n, na.rm = TRUE), .groups = "drop") |>
      mutate(
        label        = p$label,
        n_months     = n_months,
        per_month    = total / n_months
      )
  }
  
  summary_data <- bind_rows(
    summarise_period(period_1),
    summarise_period(period_2)
  ) |>
    select(series, label, total, n_months, per_month) |>
    pivot_wider(
      names_from  = label,
      values_from = c(total, n_months, per_month),
      names_glue  = "{label}_{.value}"
    ) |>
    select(series, ends_with("_per_month")) |>
    mutate(
      pct_change = round((!!sym(paste0(period_2$label, "_per_month")) -
                          !!sym(paste0(period_1$label, "_per_month"))) /
                          !!sym(paste0(period_1$label, "_per_month")) * 100, 1),
      across(ends_with("_per_month"), round, digits = 2)
    )
  
  
  return(list(p, plot_data, summary_data))
}


for (proj in unique(cd_in_projs_summed$project_name)) {  
  
  rvs <- plot_crashes_over_time(
    data             = filter(cd_in_projs_summed, year >= 2016 &
                              project_name == proj),      
    smoothing_factor = 6,
    period_1         = list(label = "Pre",  start = "2017-02-01", end = "2020-01-31"),
    period_2         = list(label = "Post", start = "2023-11-01", end = "2024-10-31"),
    title            = proj
  )
  
  print(rvs[[1]])
  print(rvs[[3]])
}
```





































<br><br>

DDOT looked at a few projects in depth. But we can also use DC's crash data to look at every protected and buffered bike lane in DDOT's "bike lane" shapefile.

what we'll do:

* get 10 meter buffer around every individual road segment
* count the number of crashes in each buffered segment in each year
  * if crashes fall into multiple buffers, assign them to a single buffer randomly. this adds noise and will attenuate the "treatment effect"
* merge on which segments had PBLs & buffered bike lanes installed

* run regression that compares crash numbers per year, controlling for area, pre-install crashes, road width, bike traffic from Cabi system, etc
* compare simple before and after means on PBL/BBL & non-PBL/BBL segments


```{r analysis_globals, fig.width=12}
# which years to include in the analysis
ANALYSIS_YEARS = c(2016, 2017, 2018, 2019, 2024, 2025)
POST_YEARS     = c(2024, 2025)
# whether to analyze the data at the subblock level (SUBBLOCKKEY) or 
# block level (BLOCKKEY):
ANALYSIS_LEVEL = "SUBBLOCKKEY"
# which treatment groups to include, options include
# switcher     : units that got PBLs between 2020 and 2023 inclusive 
# never_taker  : units that never got PBLs
# late_taker   : units that got PBLs in 2024 or 2025
# always_taker : units that got PBLs before 2019
TREATMENT_GROUPS_TO_INCLUDE = c("switcher", "never_taker")
# whether to include buffered bike lanes as well or just fully protected bike lanes
PBLS_ONLY = TRUE
# a bbox for map plotting so we don't plot tons of data at once:
PLOT_BBOX = c(xmin = -77.015, ymin = 38.94, xmax = -77.010, ymax = 38.944)

# load data:
cd                <- load_crash_data()
pbl               <- load_pbl_data(PBLS_ONLY)
cabi_sf           <- load_cabi_data()[[1]]
annual_cabi_rides <- load_cabi_data()[[2]]
rs                <- load_subblocks()

# source the R script in which we load all of the data:
source("prepare_timeseries.R")

plot_vars <- c(
  "crashes_per_acre",
  
  "injury_crashes_per_acre",
  "driver_injury_crashes_per_acre",
  # "serious_driver_injury_crashes_per_acre",
  
  "bike_injury_crashes_per_acre",
  # "serious_bike_injury_crashes_per_acre",
  
  "ped_injury_crashes_per_acre"
  # "serious_ped_injury_crashes_per_acre"
)

group_types <- c("switcher", "never_taker", "late_taker", "always_taker")

# Nicer labels for facet strip text
var_labels <- c(
  "crashes_per_acre"                      = "All Crashes",
  "injury_crashes_per_acre"               = "Injury Crashes",
  "driver_injury_crashes_per_acre"        = "Driver Injury Crashes",
  # "serious_driver_injury_crashes_per_acre"= "Serious Driver Injury Crashes",
  "bike_injury_crashes_per_acre"          = "Bike Injury Crashes",
  # "serious_bike_injury_crashes_per_acre"  = "Serious Bike Injury Crashes",
  "ped_injury_crashes_per_acre"           = "Ped Injury Crashes"
  # "serious_ped_injury_crashes_per_acre"   = "Serious Ped Injury Crashes"
)

group_labels <- c(
  "switcher"     = "Switchers (PBL installed 2020–2023)",
  "never_taker"  = "Never Takers (no PBL)",
  "late_taker"   = "Late Takers (PBL installed 2024–2025)",
  "always_taker" = "Always Takers (PBL installed before 2019)"
)


# clean the data:
res <- prepare_timeseries()
cd_pbl_clean_simplified <- res[[1]]

# descriptive histograms

# density_plots        <- map(set_names(group_types), \(g) make_density_plots(cd_pbl_clean_simplified, g))
# density_plots_log    <- map(set_names(group_types), \(g) make_density_plots(cd_pbl_clean_simplified, g, log_transform = TRUE))
# 
# walk(density_plots_log, print)
```




```{r, fig.width=12}
ecdf_plots <- map(set_names(group_types), \(g) make_ecdf_plots(cd_pbl_clean_simplified, g))
walk(ecdf_plots, print)
```

```{r}
# Usage:
summarise_outcomes_by_period(cd_pbl_clean_simplified)

# Filtered to switchers only, for example:
cd_pbl_clean_simplified |>
  filter(switcher == 1) |>
  summarise_outcomes_by_period()
```

```{r}
cd_pbl_clean_simplified |>
  filter(never_taker == 1) |>
  summarise_outcomes_by_period()
```

```{r}
cd_pbl_clean_simplified |>
  filter(always_taker == 1 | switcher == 1) |>
  summarise_outcomes_by_period()
```









































```{r, fig.width=12}


run_regs <- function(df, outcome, include_year=T, return_mod=F) {
  print(paste("############### ", outcome))
  print("################################")
  root_reg <- paste0(outcome, " ~ treatment + post_install + treatment*post_install")
  print(summary(lm(formula = as.formula(root_reg), data = df)))
  cat("\n****************************************************")
  print(summary(lm(formula = as.formula(paste0(root_reg, " + mean_crashes_pre2019")), data = df)))
  cat("\n****************************************************")
  print(summary(lm(formula = as.formula(paste0(root_reg, " + mean_crashes_pre2019", 
        ifelse(include_year, " + as.factor(year)", ""))), data = df)))
  cat("\n****************************************************")
  full_mod <- lm(formula = as.formula(
    paste0(root_reg, " + mean_crashes_pre2019 + TOTALTRAVELLANEWIDTH + TOTALPARKINGLANES ",
          "+ as.factor(WARD_ID) + as.factor(DCFUNCTIONALCLASS) + total_change_trips + cabi_rides",
           ifelse(include_year, " + as.factor(year)", ""))),
    data = df)
  print(summary(full_mod))

  if (return_mod) {
    return(full_mod)
  }
}

# run_regs(df = cd_pbl_clean, outcome = "crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "driver_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_driver_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "bike_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_bike_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "ped_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_ped_injury_crashes_per_acre")
# 
# to_plot <- run_regs(df = cd_pbl_clean, outcome = "injury_crashes_per_acre", return_mod=T)
# plot(to_plot)
```


```{r}
run_regs <- function(df, outcome, include_year = TRUE) {
  root_reg <- paste0(outcome, " ~ treatment + post_install + treatment*post_install")
  year_term <- ifelse(include_year, " + as.factor(year)", "")

  formulas <- list(
    simple = as.formula(root_reg),
    full   = as.formula(paste0(
      root_reg,
      " + mean_crashes_pre2019 + TOTALTRAVELLANEWIDTH + TOTALPARKINGLANES",
      " + as.factor(WARD_ID) + as.factor(DCFUNCTIONALCLASS) + total_change_trips",
      year_term
    ))
  )

  map_dfr(formulas, \(f) {
    tidy(lm(f, data = df)) |>
      filter(term == "treatment:post_install") |>
      mutate(outcome = outcome)
  }, .id = "model")
}


plot_effects <- function(outcomes, df) {
  results <- map_dfr(outcomes, \(o) {
    run_regs(df = df, outcome = o, include_year = FALSE)
  })
  
  results <- results |>
    mutate(significance = case_when(
        p.value < 0.001 ~ "p < 0.001",
        p.value < 0.01  ~ "p < 0.01",
        p.value < 0.05  ~ "p < 0.05",
        p.value < 0.10  ~ "p < 0.10",
        TRUE            ~ "p ≥ 0.10"
      ),
      significance = factor(significance,
                            levels = c("p < 0.001", "p < 0.01",
                                       "p < 0.05", "p < 0.10",  "p ≥ 0.10")),
      model  = factor(model, levels = c("simple", "full"),
                      labels = c("Simple model", "Full model")),
      outcome = factor(outcome, levels = outcomes) |> # preserve ordering
        fct_relabel(\(x) str_replace_all(x, "_", " ") |>
                                str_replace(" per acre", "") |>
                                str_to_sentence())
    )
  
  sig_colors <- c(
    "p < 0.001"       = "#2166ac",
    "p < 0.01"        = "#74add1",
    "p < 0.05"        = "#fdae61",
    "p < 0.10"        = "#fee08b",
    "p ≥ 0.10"        = "#d9d9d9"
  )
  
  ggplot(
          results,
         # filter(results, model=="Simple model"), 
         aes(x = outcome, y = estimate, fill = significance)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_errorbar(
      aes(ymin = estimate - 1.645 * std.error,
          ymax = estimate + 1.645 * std.error),
      position = position_dodge(width = 0.7),
      width = 0.2, color = "grey30"
    ) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
    scale_fill_manual(values = sig_colors, drop = FALSE) +
    scale_x_discrete(labels = \(x) str_replace_all(x, "_", "\n")) +
    facet_wrap(~model, ncol=1) +
    labs(
      title = "Effect of protected bike lane installation on crash outcomes",
      subtitle = "Coefficient on treatment × post-install interaction",
      x     = NULL,
      y     = "Estimated effect per acre",
      fill  = "Significance"
    ) +
    theme_minimal() +
    theme(
      axis.text.x     = element_text(size = 8),
      legend.position = "bottom",
      panel.grid.major.x = element_blank()
    )
}


plot_effects(
  outcomes = c(
      "crashes_per_acre",
      "injury_crashes_per_acre",
      
      "driver_injury_crashes_per_acre",
      "serious_driver_injury_crashes_per_acre",
  
      "bike_injury_crashes_per_acre",
      "serious_bike_injury_crashes_per_acre",
      
      "ped_injury_crashes_per_acre",
      "serious_ped_injury_crashes_per_acre"
      ),
  df = filter(cd_pbl_clean_simplified, (switcher == 1) | (never_taker == 1) | (late_taker == 1))  
  )

# plot_effects(
#   outcomes = c(
#       "total_vehicles_per_acre",
#       "total_bikes_per_acre",
#       "total_peds_per_acre"
#       )
#   )
```



```{r}
  outcomes = c(
      "crashes_per_acre",
      "injury_crashes_per_acre",
      
      "driver_injury_crashes_per_acre",
      "serious_driver_injury_crashes_per_acre",
  
      "bike_injury_crashes_per_acre",
      "serious_bike_injury_crashes_per_acre",
      
      "ped_injury_crashes_per_acre",
      "serious_ped_injury_crashes_per_acre"
      )

summary_df <-
  cd_pbl_clean_simplified |>
  filter(switcher == 1 | never_taker == 1) |>
  select(!!!sym(ANALYSIS_LEVEL), post_install, treatment, ends_with("_per_acre")) |>
  pivot_longer(cols = all_of(outcomes),
               names_to  = "outcome",
               values_to = "value") |>
  group_by(outcome, post_install, treatment) |>        # add treatment to grouping
  summarise(
    mean = mean(value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    post_install = factor(post_install, levels = c(0, 1),
                          labels = c("Pre", "Post")),
    treatment = factor(treatment, levels = c(0, 1),
                       labels = c("Other streets", "PBL installs")),
    outcome = factor(outcome, levels = outcomes) |>
                fct_relabel(\(x) str_replace_all(x, "_", " ") |>
                              str_replace(" per acre", "") |>
                              str_to_sentence())
  )

ggplot(summary_df, aes(x = post_install, y = mean,
                       color = treatment, group = treatment)) +
  geom_point(size = 3) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~outcome, nrow = 1,
             labeller = labeller(outcome = label_wrap_gen(15))) +
  scale_color_manual(values = c("Other streets"   = "grey50",
                                "PBL installs" = "#4575b4")) +
  labs(
    title = "Mean annual crashes per acre before and after bike lane installation",
    x     = NULL,
    y     = "Crashes per acre",
    color = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position    = "bottom",
    panel.grid.major.x = element_blank(),
    strip.text         = element_text(size = 7)
  )
```




```{r}

# # Get unique treated subblock keys
# treated_keys <- cd_pbl_clean_simplified |>
#   filter(treatment == 1) |>
#   distinct(SUBBLOCKKEY)
# 
# # Merge onto rs geometry
# rs_treatment <- rs |>
#   inner_join(treated_keys, by = "SUBBLOCKKEY")
# 
# # Plot
# leaflet(rs_treatment |> st_zm(drop = TRUE)) |>
#   addProviderTiles(providers$CartoDB.Positron) |>
#   addPolylines(
#     color   = "#4575b4",
#     weight  = 2.5,
#     opacity = 0.85,
#     label   = ~SUBBLOCKKEY,
#     labelOptions = labelOptions(textsize = "12px", direction = "auto"),
#     highlightOptions = highlightOptions(
#       weight      = 7,
#       color       = "#222222",
#       opacity     = 1,
#       bringToFront = TRUE
#     )
#   )

```

























