---
title: "DDOT project evaluations"
author: "pete rodrigue"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, message=F, warning=F)

setwd("~/GitHub/ddot-project-evals")

# Libraries we need:
library(readr)
library(ggplot2)
library(dplyr)
library(stringr)
library(purrr)
library(broom)
library(sf)
library(leaflet)
library(tidyr)
library(tidyverse)

# google sheet w/ the data:
url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vQtDsTt9VVyn-wptsMT6vxtCM2J1dPYWji90OOWj1v4c4pPz6Kwa6mQtOk096EU3npvJ9g4fcjnwLpp/pub?gid=1445637369&single=true&output=csv"

# load the data from the google sheet:
df <- read_csv(url, show_col_types = F)
```


```{r}
to_plot <-
  df %>%
  filter(Outcome %in% c("Injury crashes", "All crashes")) %>%
  mutate(Change = Change*-1)

ggplot(to_plot) +
  geom_bar(aes(x=`Project name`, y=Change, fill=Outcome), stat="identity", position="dodge") +
  labs(y="% decrease in crashes", x="") +
  ggtitle("All projects saw decreases in crashes") +
  geom_hline(aes(yintercept=mean(Change[Outcome == "All crashes"], na.rm=T)), color="#F8766D", linetype="dashed", linewidth=.8, alpha=.5) +
  geom_hline(aes(yintercept=mean(Change[Outcome == "Injury crashes"], na.rm=T)), color="#00BFC4", linetype="dashed", linewidth=.8, alpha=.5) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

```{r}
to_plot <-
  df %>%
  filter(Outcome %in% c("Pedestrian injury crashes")) %>%
  mutate(Change = Change*-1)

ggplot(to_plot) +
  geom_bar(aes(x=`Project name`, y=Change), stat="identity", fill="seagreen") +
  labs(y="% decrease in crashes", x="") +
  ggtitle("Pedestrian crashes fell after project completion, sometimes to zero") +
  geom_hline(aes(yintercept=mean(Change, na.rm=T)), color="grey", linetype="dashed", size=1) +
  theme_minimal() +
  theme(legend.title = element_blank())
```



```{r}
to_plot <-
  df %>%
  filter(str_detect(Outcome, "Private motor vehicle")) %>%
  group_by(`Project name`, Outcome) %>%
  summarize(change = mean(Change, na.rm=T)) %>%
  ungroup() %>%
  mutate(Outcome = ifelse(Outcome == "Private motor vehicle AM peak travel time (seconds)", 
                          "AM traffic travel time",
                          "PM traffic travel time"))

ggplot(to_plot) +
  geom_point(
    aes(x=change, color=Outcome, y=`Project name`), size=3, alpha=.7
  ) +
  geom_vline(aes(xintercept=0), color="black") +
  geom_vline(aes(xintercept=mean(change[Outcome == "AM traffic travel time"], na.rm=T)), color="#F8766D", linetype="dashed") +
  geom_vline(aes(xintercept=mean(change[Outcome == "PM traffic travel time"], na.rm=T)), color="#00BFC4", linetype="dashed") +
  labs(x="Change in travel time in seconds", y="") +
  theme_minimal() +
  xlim(c(min(to_plot$change)-10, max(to_plot$change)+25)) +
  theme(legend.title = element_blank()) +
  ggtitle("Drivers' travel times faster on average after project completion")
```



```{r}
# load crash data
# data downloaded from here:
# https://opendata.dc.gov/datasets/crashes-in-dc/about
cd <- readr::read_csv("Crashes_in_DC.csv")

# create year and month of crash variables:
cd$year  <- as.numeric(substr(x = cd$REPORTDATE, start = 1, stop = 4))
cd$month <- as.numeric(substr(x = cd$REPORTDATE, start = 6, stop = 7))

# remove rows with no date information:
cd <- filter(cd, !is.na(year))

# project crash data:
cd <- cd |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

injury_vars <- c(
  "MAJORINJURIES_BICYCLIST",   "MINORINJURIES_BICYCLIST",
  "UNKNOWNINJURIES_BICYCLIST", "FATAL_BICYCLIST",
  "MAJORINJURIES_DRIVER",      "MINORINJURIES_DRIVER",
  "UNKNOWNINJURIES_DRIVER",    "FATAL_DRIVER",
  "MAJORINJURIES_PEDESTRIAN",  "MINORINJURIES_PEDESTRIAN",
  "UNKNOWNINJURIES_PEDESTRIAN","FATAL_PEDESTRIAN"
)

cd <- select(cd, CRIMEID, year, month, injury_vars)
```
rough plan:

* get 10 meter buffer around every individual road segment
* count the number of crashes in each buffered segment in each year
  * if crashes fall into multiple buffers, assign them to a single buffer randomly. this adds noise and will attenuate the "treatment effect"
* merge on which segments had PBLs & buffered bike lanes installed

* run regression that compares crash numbers per year, controlling for area, pre-install crashes, road width, bike traffic from Cabi system, etc
* compare simple before and after means on PBL/BBL & non-PBL/BBL segments

# TODO:

* fix PBL/BBL labeling so it doesn't suck lol




```{r}

# Downloads and row binds Capital Bikeshare trip data for a given year,
# returning a data frame with all the data from that year.
#
# Data source: https://capitalbikeshare.com/system-data
# Note: CaBi publishes one zip file per month containing a CSV of individual
# trips. 
# Args: year (int): The calendar year to retrieve data for (e.g. 2023).
# Returns: A tibble with one row per trip
#
# Example: cabi_2023 <- get_annual_cabi_data(2025)
  
get_annual_cabi_data <- function(year) {

  # Build one URL per month using CaBi's S3 naming convention:
  # YYYYMM-capitalbikeshare-tripdata.zip
  urls <- sprintf(
    "https://s3.amazonaws.com/capitalbikeshare-data/%d%02d-capitalbikeshare-tripdata.zip",
    year, 1:12
  )

  trips <- map(urls, \(url) {

    # Use a unique temp file and directory per month to avoid
    # collisions when running in parallel or across multiple calls
    tmp   <- tempfile(fileext = ".zip")
    exdir <- tempfile()
    dir.create(exdir)

    tryCatch({

      download.file(url, tmp, mode = "wb", quiet = TRUE)
      unzip(tmp, exdir = exdir)

      # Some months nest CSVs inside subdirectories within the zip,
      # so use recursive = TRUE to find them wherever they land
      csv_files <- list.files(exdir, pattern = "\\.csv$",
                              full.names = TRUE, recursive = TRUE)

      # Some months ship multiple CSVs (one per week); bind them all
      map_dfr(csv_files, read_csv, show_col_types = FALSE)

    }, error = \(e) {
      # Silently skip months that don't exist yet (e.g. future months)
      # or fail to download, but surface the error message for debugging
      message("Failed to retrieve: ", url, "\n", e$message)
      NULL
    })

  }) |>
    compact() |>   # drop NULLs from failed months
    bind_rows()
  
  trips
}

# download all the trip data:
# raw_cabi_2025 <- get_annual_cabi_data(2025)
# raw_cabi_2018 <- get_annual_cabi_data(2018)

# Aggregate to station level, keeping location coordinates for
  # spatial joining downstream
# cabi_2025 <-
#   raw_cabi_2025 |>
#   filter(!is.na(end_station_id)) |>
#   group_by(end_station_id, end_lat, end_lng) |>
#   summarise(n_trips_ended = n(), .groups = "drop") |>
#   arrange(desc(n_trips_ended))

# write.csv(x = cabi_2025, file = "cabi_data_2025.csv", row.names = F)

# cabi_2018 <-
#   raw_cabi_2018 |>
#   filter(!is.na(`End station number`)) |>
#   group_by(`End station number`) |>
#   summarise(n_trips_ended = n(), .groups = "drop") |>
#   arrange(desc(n_trips_ended))

# write.csv(x = cabi_2018, file = "cabi_data_2018.csv", row.names = F)

# cabi_2018 <- readr::read_csv("cabi_data_2018.csv")
# cabi_2025 <- readr::read_csv("cabi_data_2025.csv")
# 
# # merge the 2018 and 2025 data based on the end station ID
# cabi <- 
#   dplyr::full_join(
#     cabi_2018 |>
#       rename(n_trips_ended_2018 = n_trips_ended)
#       , 
#     select(cabi_2025, end_station_id, end_lat, end_lng, n_trips_ended) |>
#       rename(n_trips_ended_2025 = n_trips_ended)
#     ,
#     by = c("End station number"="end_station_id")
#     )
# 
# add_coords <- function(df, station_id, lat_lon) {
#   
#   df$end_lat[df$`End station number`==station_id] <- lat_lon[1]
#   df$end_lng[df$`End station number`==station_id] <- lat_lon[2]
#   
#   df
# }
# # add lat lon for 2018 stations that were later removed but had at least
# # 1000 trips end at that station in 2018
# cabi <- add_coords(cabi, 31613, c(38.88429087871622, -76.99578835580621))
# cabi <- add_coords(cabi, 31614, c(38.900116296819704, -76.99149861721462))
# cabi <- add_coords(cabi, 31103, c(38.926263953061635, -77.03661594753648))
# cabi <- add_coords(cabi, 31008, c(38.8629439551263, -77.05208676187382))
# 
# write.csv(cabi, "cabi_both_years_clean.csv", row.names = F)

cabi <- readr::read_csv("cabi_both_years_clean.csv")

cabi <-
  cabi |>
  mutate(
    n_trips_ended_2018 = if_else(is.na(n_trips_ended_2018), 0, n_trips_ended_2018),
    n_trips_ended_2025 = if_else(is.na(n_trips_ended_2025), 0, n_trips_ended_2025),
    change_n_trips     = n_trips_ended_2025 - n_trips_ended_2018
  ) %>%
  filter(!is.na(end_lat))



# Diverging palette centered at 0 — red = fewer trips, blue = more trips
pal <- colorNumeric(
  palette = "RdBu",
  domain  = cabi$change_n_trips,
  reverse = FALSE
)

leaflet(cabi) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addCircleMarkers(
    lng          = ~end_lng,
    lat          = ~end_lat,
    radius       = 6,
    color        = "white",
    weight       = 0.5,
    fillColor    = ~pal(change_n_trips),
    fillOpacity  = 0.85,
    label = ~paste0(
      "<b>", `End station number`, "</b><br>",
      "2018 trips: ", scales::comma(n_trips_ended_2018), "<br>",
      "2025 trips: ", scales::comma(n_trips_ended_2025), "<br>",
      "Change: ", scales::comma(change_n_trips)
    ) |> lapply(htmltools::HTML),
    labelOptions = labelOptions(textsize = "12px", direction = "auto")
  ) |>
  addLegend(
    position = "bottomright",
    pal      = pal,
    values   = ~change_n_trips,
    title    = "Change in<br>annual trips",
    labFormat = labelFormat(big.mark = ","),
    opacity  = 0.9
  ) |>
  addScaleBar(position = "bottomleft")

# Convert cabi to sf object
cabi_sf <- cabi |>
  filter(!is.na(end_lat), !is.na(end_lng)) |>
  st_as_sf(coords = c("end_lng", "end_lat"), crs = 4326) |>
  st_transform(st_crs(4326))


```

```{r}
# load labeled PBL data that has the road subblock IDs and the year in which 
# PBLs were installed on those subblocks
pbl <- st_read("bike_lanes_labeled.geojson") %>%
  filter(!is.na(build_year)) %>%
  # ugh gotta add back missing subblocks wtf qgis
  add_row(SUBBLOCKKE="22ef82997ee696dd668ca7f21c7ab4ac", build_year=2023, ROUTENAME="39TH ST NW") |>
  add_row(SUBBLOCKKE="1175ddbdaed945393b93d8501a034559", build_year=2023, ROUTENAME="39TH ST NW") |>
  add_row(SUBBLOCKKE="a914485d49b99b00499ac25f4cfa3e55", build_year=2023, ROUTENAME="39TH ST NW") #|>
  # add_row(SUBBLOCKKE="22ef82997ee696dd668ca7f21c7ab4ac", build_year=2023, ROUTENAME="blank") |>
  # add_row(SUBBLOCKKE="22ef82997ee696dd668ca7f21c7ab4ac", build_year=2023, ROUTENAME="blank") |>
  # add_row(SUBBLOCKKE="22ef82997ee696dd668ca7f21c7ab4ac", build_year=2023, ROUTENAME="blank") |>
  # add_row(SUBBLOCKKE="22ef82997ee696dd668ca7f21c7ab4ac", build_year=2023, ROUTENAME="blank") |>


# load all roadway sub-block data, from here:
# https://opendata.dc.gov/datasets/df571ab7fea446e396bf2862d0ab6833_162/explore?location=38.927746%2C-77.024575%2C18
rs <- 
  st_read("Roadway_SubBlock.geojson") %>%
  st_zm(drop = TRUE, what = "ZM") %>%  # drops Z (and M if present) → pure XY
  # Leaflet requires WGS84 (EPSG:4326) — reproject if needed
  st_transform(4326) %>%
  st_cast("MULTILINESTRING") 


rs_cabi_buf <-
  rs |>
  st_transform(6487) |>          # EPSG:6487 — NAD83 / Maryland (metres)
  st_buffer(500) |>               # X m buffer
  st_transform(st_crs(cd))       # reproject back to match cd's CRS

# Spatial join stations to buffers, then summarise by subblock
rs_cabi_change <- rs_cabi_buf |>
  st_join(cabi_sf, join = st_contains) |>
  st_drop_geometry() |>
  group_by(SUBBLOCKKEY) |>
  summarise(
    total_change_trips = sum(change_n_trips, na.rm = TRUE),
    n_stations         = sum(!is.na(change_n_trips)),
    .groups = "drop"
  )

rs_crash_buf <-
  rs |>
  st_transform(6487) |>          # EPSG:6487 — NAD83 / Maryland (metres)
  st_buffer(10) |>               # X m buffer
  st_transform(st_crs(cd)) |>    # reproject back to match cd's CRS
  select(ROUTEID, ROUTENAME, BLOCKKEY, SUBBLOCKKEY, 
         TOTALTRAVELLANES, TOTALPARKINGLANES, TOTALTRAVELLANEWIDTH,
         WARD_ID, DCFUNCTIONALCLASS
         ) %>%
  mutate(area_m = st_area(.))

units(rs_crash_buf$area_m) <- NULL
# Join crashes to buffers — one crash can match multiple buffers
cd_joined <- cd |> st_join(rs_crash_buf, join = st_within, left = FALSE)

# For crashes in multiple buffers, randomly keep one
cd_assigned <- cd_joined |>
  st_drop_geometry() |>
  group_by(CRIMEID) |>           # replace crash_id with whatever your unique crash ID column is
  slice_sample(n = 1) |>         # randomly pick one buffer per crash
  ungroup()



cd_assigned_clean <-
  cd_assigned %>%
  mutate(WARD_ID = as.numeric(WARD_ID)) %>%
  group_by(SUBBLOCKKEY, year) %>%
  summarise(
    across(all_of(injury_vars), \(x) sum(x, na.rm = TRUE)),
    n_crashes            = n(),
    TOTALTRAVELLANES     = first(TOTALTRAVELLANES),
    TOTALPARKINGLANES    = first(TOTALPARKINGLANES),
    TOTALTRAVELLANEWIDTH = first(TOTALTRAVELLANEWIDTH),
    WARD_ID              = first(WARD_ID),
    DCFUNCTIONALCLASS    = first(DCFUNCTIONALCLASS),
    area_m               = mean(area_m, na.rm=T),
    .groups   = "drop"
  ) %>%
  ungroup() %>%
  mutate(
    crashes_per_acre     = n_crashes / area_m * 4046.86
  ) %>%
  mutate(across(where(is.numeric), ~ ifelse(is.nan(.), NA, .)))


cd_assigned_clean <- cd_assigned_clean |>
  complete(
    SUBBLOCKKEY,
    year = 2014:2025,
    fill = list(
      n_crashes            = 0,
      TOTALTRAVELLANES     = NA_real_,
      TOTALPARKINGLANES    = NA_real_,
      TOTALTRAVELLANEWIDTH = NA_real_,
      WARD_ID              = NA_real_,
      area_m               = NA_real_,
      DCFUNCTIONALCLASS    = NA_real_,
      crashes_per_acre     = 0,
      # injury vars
      MAJORINJURIES_BICYCLIST   = 0,
      MINORINJURIES_BICYCLIST   = 0,
      UNKNOWNINJURIES_BICYCLIST = 0,
      FATAL_BICYCLIST           = 0,
      MAJORINJURIES_DRIVER      = 0,
      MINORINJURIES_DRIVER      = 0,
      UNKNOWNINJURIES_DRIVER    = 0,
      FATAL_DRIVER              = 0,
      MAJORINJURIES_PEDESTRIAN  = 0,
      MINORINJURIES_PEDESTRIAN  = 0,
      UNKNOWNINJURIES_PEDESTRIAN= 0,
      FATAL_PEDESTRIAN          = 0
    )
  )

cd_assigned_clean <- cd_assigned_clean |>
  arrange(SUBBLOCKKEY, year) |>
  group_by(SUBBLOCKKEY) |>
  fill(TOTALTRAVELLANES, TOTALPARKINGLANES, TOTALTRAVELLANEWIDTH, DCFUNCTIONALCLASS,
       WARD_ID, area_m, .direction = "downup") |>
  ungroup()

cd_pbl_clean <- 
  dplyr::left_join(cd_assigned_clean,
                   pbl %>% 
                     st_drop_geometry() %>%
                     select(SUBBLOCKKE, build_year) %>%
                     group_by(SUBBLOCKKE) %>%
                     summarise(
                       min_build_year = min(build_year, na.rm=T),
                       max_build_year = max(build_year, na.rm=T)
                               ) %>%
                     ungroup() %>%
                     rename(SUBBLOCKKEY = SUBBLOCKKE),
                   by="SUBBLOCKKEY"
                   )

cd_pbl_clean <- cd_pbl_clean %>%
  filter(year %in% c(2014, 2015, 2016, 2017, 2018, 2024, 2025)) %>%
  filter((max_build_year <= 2023) | is.na(max_build_year)) %>%
  mutate(post_install = if_else(year > 2023, 1, 0)) %>%
  group_by(SUBBLOCKKEY) %>%
  mutate(
    min_post_val = min(post_install, na.rm=T),
    max_post_val = max(post_install, na.rm=T)
         ) %>%
  ungroup() %>%
    filter((min_post_val==0) & (max_post_val==1)) %>%
  mutate(
    n_injury_crashes = 
           MAJORINJURIES_BICYCLIST + MINORINJURIES_BICYCLIST + UNKNOWNINJURIES_BICYCLIST + FATAL_BICYCLIST +
           MAJORINJURIES_DRIVER + MINORINJURIES_DRIVER + UNKNOWNINJURIES_DRIVER + FATAL_DRIVER +
           MAJORINJURIES_PEDESTRIAN + MINORINJURIES_PEDESTRIAN + UNKNOWNINJURIES_PEDESTRIAN + FATAL_PEDESTRIAN,
    n_driver_injury_crashes = 
           MAJORINJURIES_DRIVER + MINORINJURIES_DRIVER + UNKNOWNINJURIES_DRIVER + FATAL_DRIVER,
    n_serious_driver_injury_crashes = 
           MAJORINJURIES_DRIVER + FATAL_DRIVER,
    n_bike_injury_crashes = 
           MAJORINJURIES_BICYCLIST + MINORINJURIES_BICYCLIST + UNKNOWNINJURIES_BICYCLIST + FATAL_BICYCLIST,
    n_serious_bike_injury_crashes = 
           MAJORINJURIES_BICYCLIST + FATAL_BICYCLIST,
    n_ped_injury_crashes = 
           MAJORINJURIES_PEDESTRIAN + MINORINJURIES_PEDESTRIAN + UNKNOWNINJURIES_PEDESTRIAN + FATAL_PEDESTRIAN,
    n_serious_ped_injury_crashes = 
           MAJORINJURIES_PEDESTRIAN + FATAL_PEDESTRIAN,
    
    injury_crashes_per_acre                = n_injury_crashes / area_m * 4046.86,
    driver_injury_crashes_per_acre         = n_driver_injury_crashes / area_m * 4046.86,
    serious_driver_injury_crashes_per_acre = n_serious_driver_injury_crashes / area_m * 4046.86,
    bike_injury_crashes_per_acre           = n_bike_injury_crashes / area_m * 4046.86,
    serious_bike_injury_crashes_per_acre   = n_serious_bike_injury_crashes / area_m * 4046.86,
    ped_injury_crashes_per_acre            = n_ped_injury_crashes / area_m  * 4046.86,
    serious_ped_injury_crashes_per_acre    = n_serious_ped_injury_crashes / area_m  * 4046.86,
    
    treatment = if_else(is.na(max_build_year), 0, 1)
           ) %>%
  group_by(SUBBLOCKKEY, post_install) %>%
  mutate(
    mean_crashes_pre2019 = mean(crashes_per_acre[year < 2019], na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(mean_crashes_pre2019 = na_if(mean_crashes_pre2019, NaN)) %>%
  group_by(SUBBLOCKKEY) %>%
  mutate(mean_crashes_pre2019 = mean(mean_crashes_pre2019, na.rm=T)) %>%
  ungroup()



cd_pbl_clean <- dplyr::left_join(cd_pbl_clean, rs_cabi_change, by="SUBBLOCKKEY")

# run_regs <- function(df, outcome, include_year=T) {
#   root_reg <- paste0(outcome, " ~ treatment + post_install + treatment*post_install")
#   print(summary(lm(formula = as.formula(root_reg), data = df)))
#   cat("\n****************************************************")
#   print(summary(lm(formula = as.formula(paste0(root_reg, " + mean_crashes_pre2019")), data = df)))
#   cat("\n****************************************************")
#   print(summary(lm(formula = as.formula(paste0(root_reg, " + mean_crashes_pre2019", ifelse(include_year, " + as.factor(year)", ""))), data = df)))
#   cat("\n****************************************************")
#   print(summary(lm(formula = as.formula(
#     paste0(root_reg, " + mean_crashes_pre2019 + TOTALTRAVELLANEWIDTH + TOTALPARKINGLANES + as.factor(WARD_ID) + as.factor(DCFUNCTIONALCLASS) + total_change_trips",
#            ifelse(include_year, " + as.factor(year)", ""))), 
#     data = df)))
# }

# run_regs(df = cd_pbl_clean, outcome = "crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "driver_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_driver_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "bike_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_bike_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "ped_injury_crashes_per_acre")
# run_regs(df = cd_pbl_clean, outcome = "serious_ped_injury_crashes_per_acre")


cd_pbl_clean_simplified <-
  cd_pbl_clean %>%
  group_by(SUBBLOCKKEY, post_install) %>%
  summarize(
    TOTALTRAVELLANEWIDTH = mean(TOTALTRAVELLANEWIDTH, na.rm=T),
    TOTALPARKINGLANES    = mean(TOTALPARKINGLANES, na.rm=T),
    DCFUNCTIONALCLASS    = mean(DCFUNCTIONALCLASS, na.rm=T),
    WARD_ID              = mean(WARD_ID, na.rm=T),
    total_change_trips   = mean(total_change_trips, na.rm=T),
    
    across(ends_with("_per_acre"), \(x) mean(x, na.rm = TRUE)),
    
    treatment                            = mean(treatment, na.rm=T),
    mean_crashes_pre2019                 = mean(mean_crashes_pre2019, na.rm=T)
            ) %>%
  ungroup()





# run_regs(df = cd_pbl_clean_simplified, outcome = "crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "driver_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "serious_driver_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "bike_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "serious_bike_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "ped_injury_crashes_per_acre", F)
# run_regs(df = cd_pbl_clean_simplified, outcome = "serious_ped_injury_crashes_per_acre", F)
```


```{r}
run_regs <- function(df, outcome, include_year = TRUE) {
  root_reg <- paste0(outcome, " ~ treatment + post_install + treatment*post_install")
  year_term <- ifelse(include_year, " + as.factor(year)", "")

  formulas <- list(
    simple = as.formula(root_reg),
    full   = as.formula(paste0(
      root_reg,
      " + mean_crashes_pre2019 + TOTALTRAVELLANEWIDTH + TOTALPARKINGLANES",
      " + as.factor(WARD_ID) + as.factor(DCFUNCTIONALCLASS) + total_change_trips",
      year_term
    ))
  )

  map_dfr(formulas, \(f) {
    tidy(lm(f, data = df)) |>
      filter(term == "treatment:post_install") |>
      mutate(outcome = outcome)
  }, .id = "model")
}


outcomes <- c(
  "crashes_per_acre",
  "injury_crashes_per_acre",
  "driver_injury_crashes_per_acre",
  "serious_driver_injury_crashes_per_acre",
  "bike_injury_crashes_per_acre",
  "serious_bike_injury_crashes_per_acre",
  "ped_injury_crashes_per_acre",
  "serious_ped_injury_crashes_per_acre"
)

results <- map_dfr(outcomes, \(o) {
  run_regs(df = cd_pbl_clean_simplified, outcome = o, include_year = FALSE)
})

results <- results |>
  mutate(
    significance = case_when(
     p.value < 0.05 ~ "p < 0.05",
      p.value < 0.10 ~ "p < 0.10",
      p.value < 0.15 ~ "p < 0.15",
      TRUE           ~ "p ≥ 0.15"
    ),
    significance = factor(significance,
                          levels = c("p < 0.05", "p < 0.10",
                                     "p < 0.15", "p ≥ 0.15")),
    model  = factor(model, levels = c("simple", "full"),
                    labels = c("Simple model", "Kitchen sink model")),
    outcome = factor(outcome, levels = outcomes)  # preserve ordering
  )

sig_colors <- c(
  "p < 0.05" = "#2166ac",
  "p < 0.10" = "#74add1",
  "p < 0.15" = "#fdae61",
  "p ≥ 0.15" = "#d9d9d9"
)

```

```{r}
ggplot(filter(results, model=="Simple model"), aes(x = outcome, y = estimate, fill = significance)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(
    aes(ymin = estimate - 1.96 * std.error,
        ymax = estimate + 1.96 * std.error),
    position = position_dodge(width = 0.7),
    width = 0.2, color = "grey30"
  ) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  scale_fill_manual(values = sig_colors) +
  scale_x_discrete(labels = \(x) str_replace_all(x, "_", "\n")) +
  # facet_wrap(~model) +
  labs(
    title = "Effect of protected bike lane installation on crash outcomes",
    subtitle = "Coefficient on treatment × post-install interaction",
    x     = NULL,
    y     = "Annual crashes / acre",
    fill  = "Significance"
  ) +
  theme_minimal() +
  theme(
    axis.text.x     = element_text(size = 8),
    legend.position = "bottom",
    panel.grid.major.x = element_blank()
  ) 

```



```{r}
summary_df <-
  cd_pbl_clean_simplified |>
  select(SUBBLOCKKEY, post_install, treatment, ends_with("_per_acre")) |>
  pivot_longer(cols = all_of(outcomes),
               names_to  = "outcome",
               values_to = "value") |>
  group_by(outcome, post_install, treatment) |>        # add treatment to grouping
  summarise(
    mean = mean(value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    post_install = factor(post_install, levels = c(0, 1),
                          labels = c("Pre", "Post")),
    treatment = factor(treatment, levels = c(0, 1),
                       labels = c("Other streets", "PBL installs")),
    outcome = factor(outcome, levels = outcomes) |>
                fct_relabel(\(x) str_replace_all(x, "_", " ") |>
                              str_replace(" per acre", "") |>
                              str_to_sentence())
  )

ggplot(summary_df, aes(x = post_install, y = mean,
                       color = treatment, group = treatment)) +
  geom_point(size = 3) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~outcome, nrow = 1,
             labeller = labeller(outcome = label_wrap_gen(15))) +
  scale_color_manual(values = c("Other streets"   = "grey50",
                                "PBL installs" = "#4575b4")) +
  labs(
    title = "Mean annual crashes per acre before and after bike lane installation",
    x     = NULL,
    y     = "Crashes per acre",
    color = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position    = "bottom",
    panel.grid.major.x = element_blank(),
    strip.text         = element_text(size = 7)
  )
```




```{r}

# # Get unique treated subblock keys
# treated_keys <- cd_pbl_clean_simplified |>
#   filter(treatment == 1) |>
#   distinct(SUBBLOCKKEY)
# 
# # Merge onto rs geometry
# rs_treatment <- rs |>
#   inner_join(treated_keys, by = "SUBBLOCKKEY")
# 
# # Plot
# leaflet(rs_treatment |> st_zm(drop = TRUE)) |>
#   addProviderTiles(providers$CartoDB.Positron) |>
#   addPolylines(
#     color   = "#4575b4",
#     weight  = 2.5,
#     opacity = 0.85,
#     label   = ~SUBBLOCKKEY,
#     labelOptions = labelOptions(textsize = "12px", direction = "auto"),
#     highlightOptions = highlightOptions(
#       weight      = 7,
#       color       = "#222222",
#       opacity     = 1,
#       bringToFront = TRUE
#     )
#   )

```

























